Project assignment - Tahoe-LAFS with SHA-3 candidates

### 2010-11-16 paper/main.tex
Started Distutils section, added Cython implementation details

\begin{acronym}
\acro{AES}{Advanced Encryption Standard}
\acro{API}{Application programming interface}
\acro{CEB}{Capability Extension Block}
\acro{FEC}{Forward Error Correction}
\acro{GCC}{GNU Compiler Collection}
\acro{KAT}{Known Answer Test}
\acro{LAFS}{Least-Authority Filesystem}
\acro{NIST}{National Institute of Standards and Technology}
\acro{RAM}{Random Access Memory}
\acro{RAID}{Redundant Array of Independent Disks}
\acro{SHA}{Secure Hash Algorithm}
\acro{SSE2}{Streaming SIMD Extensions 2}
\acro{SSSE3}{Supplemental Streaming SIMD Extensions 3}
\acro{SUPERCOP}{System for Unified Performance Evaluation Related to
Cryptographic Operations and Primitives}
\acro{UEB}{URI Extension Block}
\end{acronym}

@@

Since Tahoe-\ac{LAFS} is written in Python, we needed some way of interacting
with the C implementations of the NIST candidates.

>>>>

\begin{acronym}
\acro{AES}{Advanced Encryption Standard}
\acro{API}{Application programming interface}
\acro{CEB}{Capability Extension Block}
\acro{Distutils}{Python Distribution Utilities}
\acro{FEC}{Forward Error Correction}
\acro{GCC}{GNU Compiler Collection}
\acro{KAT}{Known Answer Test}
\acro{LAFS}{Least-Authority Filesystem}
\acro{NIST}{National Institute of Standards and Technology}
\acro{RAM}{Random Access Memory}
\acro{RAID}{Redundant Array of Independent Disks}
\acro{SHA}{Secure Hash Algorithm}
\acro{SSE2}{Streaming SIMD Extensions 2}
\acro{SSSE3}{Supplemental Streaming SIMD Extensions 3}
\acro{SUPERCOP}{System for Unified Performance Evaluation Related to
Cryptographic Operations and Primitives}
\acro{UEB}{URI Extension Block}
\end{acronym}

@@

Since Tahoe-\ac{LAFS} is written in Python, we needed some way of interacting
with the C implementations of the NIST candidates. We did this by the use of
Cython, and built a python object-oriented library which we named Sha3lib.
Sha3lib supports all of the 14 second round \ac{SHA}-3 candidates, but only in
their 256-bit form.

\subsection{Sha3lib}
Sha3lib is built to mimic the functionality of
hashlib\footnote{\url{http://docs.python.org/library/hashlib.html}}, the
default python library for MD5, \ac{SHA}-1 and the \ac{SHA}-2 family. 
% TODO, figure or listing of what this means

\subsubsection{Cython}
We employ Cython to generate bindings for the C implementation and the
different candidates, as well as to create the base python classes. To access
functions in C we need to generate a Cython header file, which pretty much does
the same as a C header file; it tells which functions and data types are
available. How we accomplish this for an implementation that satisfies the
\ac{NIST} \ac{API} is shown in Listing \ref{lst:cython_header}.

\begin{lstlisting}[label=lst:cython_header, caption=Cython header\, echo\_hash\_h.pxd]
cdef extern from "sha3nist.h":
    ctypedef unsigned char BitSequence
    ctypedef unsigned long long DataLength

    cdef enum HashReturn:
        SUCCESS = 0
        FAIL = 1
        BAD_HASHLEN = 2

    ctypedef struct hashState:
        pass

    HashReturn Hash(int, BitSequence *data, DataLength, BitSequence *hashval)
    HashReturn Init(hashState *state, int hashbitlen)
    HashReturn Update(hashState *state, BitSequence *data, DataLength databitlen)
    HashReturn Final(hashState *state, BitSequence *hashval)
\end{lstlisting}

We then create a wrapper-class that mimics hashlib, this is shown in Listing
\ref{lst:cython_impl}. Functions declared with def er python functions, cdef are
Cython only while cpdef are functions that can be accessed from both. While
this is just the implementation for the echo candidate, the almost exact same
pattern is used for all of the other candidates.

\begin{lstlisting}[label=lst:cython_impl, caption=Cython class\, echo\_hash.pyx]
cimport echo_hash_h
cimport cython

from libc.stdlib cimport *

cdef class echo:
    '''
    A class that tries to mimic the behaviour of hashlib, i.e. keeping state
    so that one can update the hashing procedure instead of doing it from
    scratch.
    '''
    #cdef echo_hash_h.BitSequence *hashval
    cdef echo_hash_h.hashState previous_state
    cdef echo_hash_h.hashState state
    cdef int finished
    cdef int hashbitlen
    cdef list hashval

    def __init__(self, int in_hashbitlen, bytes initial=None):
        self.finished = 0
        self.hashbitlen = in_hashbitlen
        echo_hash_h.Init(&self.state, self.hashbitlen)

        if initial:
            self.update(initial)

    cpdef update(self, bytes in_data):
        cdef char* data = <char *> in_data
        cdef int data_len = len(in_data)*8

        if self.finished:
            self.finished = 0
            self.state = self.previous_state

        echo_hash_h.Update(&self.state, <echo_hash_h.BitSequence *> data, data_len)

    cpdef final(self):
        cdef echo_hash_h.BitSequence *hashval = <echo_hash_h.BitSequence *> malloc(self.hashbitlen*8)

        # We copy the state so that we can continue to update.
        # This equals hashlibs functionality, but not pycryptopp.
        self.previous_state = self.state

        echo_hash_h.Final(&self.state, hashval)
        self.finished = 1

        self.hashval = [hashval[i] for i from 0 <= i < self.hashbitlen / 8]
        free(hashval)

    cpdef copy(self):
        s = echo(self.hashbitlen)
        s.state = self.state
        s.previous_state = self.previous_state
        s.finished = self.finished
        if s.finished:
            s.hashval = self.hashval

        return s

    def digest(self):
        if not self.finished:
            self.final()

        return ''.join(map(chr, self.hashval))

    def hexdigest(self):
        if not self.finished:
            self.final()

        return ''.join(['%02x' % i for i in self.hashval])
\end{lstlisting}

\subsubsection{\ac{Distutils}} We use the
\ac{Distutils}\footnote{\url{http://docs.python.org/distutils/}} to build and
install our code. The alternative to doing this would be to manually transform
cython files to C-files, build the shared library from this and other relevant
C-files and then install by copying into the correct system directory where
python libraries should be stored. \ac{Distutils} also support expansion
through self defined commands, something we have utilised to do Unit Testing.
Options and configuration for \ac{Distutils} is done through a file setup.py.
Relevant configuration for compiling the echo extension can be seen in Listing.

\begin{lstlisting}[label=lst:setup_py, caption=Compiling echo from \ac{Distutils}]
from distutils.core import setup, Command
from distutils.extension import Extension
from Cython.Distutils import build_ext

srcpath = 'sha3lib/'
hashpath = srcpath + 'hash_functions/'

ext_modules = [
    Extension("sha3lib.hash_functions.echo.echo_hash",
        [hashpath+"echo/echo_hash.pyx", 
         hashpath+'echo/sha3nist.c', 
         hashpath+"echo/echo.c"],
    extra_compile_args=[
        '-funroll-loops',
        '-m32','-march=prescott',
        '-O2',
        '-fomit-frame-pointer']
    ),
]

setup(
    name='SHA3lib',
    version='0.1',
    description='Python bindings for SHA3-256 using Cython',
    cmdclass = {'build_ext': build_ext},
    ext_modules = ext_modules,
    packages =
        ['sha3lib',
        'sha3lib.hash_functions',
        'sha3lib.hash_functions.echo',
        ],
)

\end{lstlisting}
